Extracting Information from AGN Variability

### 2016-07-14 Manuscript.tex
Made Richards & Moreno edits

% Abstract of the paper
\begin{abstract}
AGN exhibit rapid, high amplitude stochastic flux variability across the entire electromagnetic spectrum on timescales ranging from hours to years. The cause of this variability is poorly understood because of the complexity of the phenomenon being observed. We present a new method for using variability to (1) measure the time-scales on which flux perturbations evolve and (2) characterize the driving flux perturbations. We model the observed light curve of an AGN as a linear differential equation driven by stochastic impulses. Physically, the impulses could be local `hot-spots` in the accretion disk---the linear differential equation then governs how the hot spots evolve and dissipate. The impulse-response function of the accretion disk material is given by the Green's function of the linear differential equation. The timescales on which the hot-spots radiate energy is characterized by the Power Spectral Density (PSD) of the driving stochastic impulses. We apply our technique for analyzing light curves to the light curve of the \Kepler AGN Zw 229-15 and find that the observed variability behavior can be modeled as a damped harmonic oscillator perturbed by a colored noise process. The model PSD turns over on time-scale $385$~d. On shorter time-scales, the model PSD log slope varies between $2$ and $4$, explaining the behavior noted by previous studies. We recover and identify both the $5.6$~d and $67$~d timescales reported by previous work. In our model, these timescales represent the time-scale on which flux perturbations grow, and the time-scale on which flux perturbations decay back to the steady-state flux level respectively. The PSD of the driving impulses is flat on long time-scales. On time-scales shorter than $1$~d, the impulse PSD begins to rise steeply possibly due to turbulent eddies in the accretion disk, or by the thermal motion of the accretion disk plasma. We have developed and tested the software package \textsc{k\={a}l\={i}} to study light curves using our methods. \textsc{k\={a}l\={i}} is freely available on Github for use by the astronomical community.

We have developed and tested a fast software package, \textsc{k\={a}l\={i}}, to analyze light curves using our method that we have made available . 
%We decompose the observed light curve fluctuations into stochastic impulses that drive an ordinary linear differential equation. The Green's function of the differential equation quantifies the time-domain evolution of perturbations to the emitted flux and hence probes dissipative physical processes in accretion flow. The PSD of the stochastic driving impulses characterizes the physical processes responsible for the flux perturbations. 
\end{abstract}

@@

Numerous mechanisms have been posited for producing variability in accretion disks \citep{DoneAccretion,MaccaroneAccretion, UttleyAccretion}. While we are concerned primarily with accretion onto supermassive black holes (SMBH) in AGN, a fair fraction of the theory is developed with BHXRB accretion in mind. While similarities exist between accretion in BHXRBs and AGN, the difference in energy densities is considerable with AGNs having considerably higher luminosities and exhibiting variability on longer time-scales. For instance, quasi-periodic oscillations (QPO) are routinely observed in BHXRBs, whereas in AGN there few reports of QPOs \citep{GierlinskiQPO,GrindleyQPO}. While BHXRBs are commonly observed to shift between various spectral states, such transitions have not been observed in AGN to date \citep{Kelly11}.

AGN optical variability may arise due to fluctuations in the \citet{ShakuraSunyaev73} viscosity parameter $\alpha$. \citet{Lyubarskii97} showed that local stochastic fluctuations in the accretion disk are capable of producing the power-law power spectral densities (PSD) observed in AGN and BHXRBs. Such fluctuations are thought to propagate inwards over time suggesting that a powerful observational test is the observation of blue-lags \citep{UttleyAccretion} i.e. light curve variations at shorter wavelengths lagging those at longer wavelengths. Evidence for the propagating fluctuation model was provided by \citet{Miyamoto88} who observed that hard X-ray variations in Cyg X-1 lagged behind soft X-ray variations with an inverse scaling between the time lag and the variability time-scale. \citet{Starling04} used the viscosity fluctuation model to impose lower limit of $\alpha \sim 0.01$ on the disk viscosity. A later study by \citet{King07} using AGN variability triggered by viscosity fluctuations suggest that $\alpha \sim 0.1$-$0.4$. \citet{Wood01} examined a variant of the \citet{Lyubarskii97} propagating fluctuations model in which episodic mass deposition at the outer edge of the accretion disk triggers viscosity fluctuations in the accretion disk. This model was revisited by \citet{Titarchuk07} who generalized the treatment of \citet{Wood01} to include time-dependent fluctuation sources located randomly on the disk face. \citet{Titarchuk07} follow the evolution of perturbations to the mass surface density and luminosity arising from viscosity that is a power-law function of disk radius. Subsequently, \citet{Kelly11} studied solutions to the surface mass density equation perturbed by stochastic fluctuations and showed that the resulting light curve can be expressed as a linear combination of multiple Ornstein-Uhlenbeck processes \citep{Gillespie96,Kelly09}.

@@

Studies of AGN variability model the observed light curve as stochastic processes i.e mathematically the light curve as a function of time is a random variable that correlates with previous values of the same light curve. Prescriptions for stochastic models require two properties to be specified: 1. the distribution from which to draw the random values of the light curve; and 2. the strength of the correlation of a given value of the light curve with previous values of the same light curve. Most studies of AGN variability (implicitly) invoke some form of the central limit theorem to argue that property 1 should be a Gaussian distribution which then lets them specify property 2 using a second-order statistic such as the power spectral density (PSD), or the auto-covariance function (ACVF), or the structure function (SF). The most straightforward method of fitting such a stochastic model to a light curve is by direct estimation of the PSD of the observed light curve. Using periodogram estimators of the PSD, the observed PSD can be compared to the predicted theoretical PSD as advocated in \citet*{Uttley02}. Alternatively, it is also possible to estimate the ACVF of the observed light curve and fit it to the theoretical ACVF of a given stochastic process. Equivalently, following \citet*{Kasliwal15}, estimates of the structure function (SF) of the observed light curve can be fit to the theoretical SF computed either directly from the theoretical ACVF (via $SF_{\chi\chi}(T) \sim -2\gamma_{\chi\chi}(T)$), or from Monte-Carlo simulations of light curves generated using the prescribed model PSD.

@@

Recently, \citet{Kelly14} suggested that stochastic variability in a large class of astrophysical sources may be modelled using Continuous-time AutoRegressive Moving Average (C-ARMA) processes. The attractiveness of using C-ARMA processes to model stochastic phenomenon stems from two factors: 1. C-ARMA processes directly specify property 2, i.e. the correlation structure of the light curve by providing a differential equation that governs how the light curve evolves over time. This alternative to providing a second-order statistic is powerful because it makes the light curve evolution more amenable to physical interpretation as is shown in this work. It also provides some idea of how to replace Gaussian distributions with more generalized $\alpha$-stable distributions that may be truer to reality \citep{BrockwellMarquardt05, BrockwellLinder09}; and 2. flexibility of the shape of their PSD. Since the PSD of a C-ARMA process is a rational function, it can be made arbitrarily complex to model a very wide range of phenomenon. This flexibility is achieved by allowing C-ARMA processes to have multiple Autoregressive (quantified by $p$) and Moving Average (quantified by $q$) terms under the constraint $q < p$. If $\chi$ is a small flux perturbation from the mean flux level, then a C-ARMA($p$,$q$) process that models the time evolution of $\chi$ has the general form
\begin{multline}\label{eq:CARMAIntro}
\mathrm{d}^{p}\chi + \alpha_{1} \mathrm{d}^{p-1}\chi + \ldots + \alpha_{p-1} \mathrm{d}\chi + \alpha_{p} \chi = \\ \beta_{0} (\mathrm{d}W) + \beta_{1} \mathrm{d}(\mathrm{d}W) + \ldots + \beta_{p-1} \mathrm{d}^{p-1}(\mathrm{d}W)
\end{multline}
where $W$ is a \textit{Wiener} process and is responsible for introducing stochasticity into the behavior of $\chi$. The $\alpha$ and $\beta$ are constants that govern the behavior of the resulting process and have dimensions of various powers of time. In this paper, we demonstrate how C-ARMA processes can arise from the physical processes in accretion disks and present a Green's function-based method for understanding what a C-ARMA process tells us about accretion physics.

@@

Appendices~\ref{sec:Fitting} through~\ref{sec:ModelSelection} present details of inferring the properties of the C-ARMA process from noisy observations using the Kalman filter. Alternatives exist to using the Kalman filter for inferring the parameters of a C-ARMA process. We discuss why the Kalman filter is the optimal choice of fitting techniques in appendix~\ref{sec:Fitting}. 
%In brief, the Kalman filter is both statistically optimal and computationally efficient. In order to use the Kalman filter, it is necessary to cast the C-ARMA process into \textit{state-space} form. 
and suggest an alternative to the traditional \textit{controllable canonical form} state-space representation used by \citet{JonesAckerson90} and \citet{Brockwell14}. Our choice of state-space representations is (1) better suited for extension to more complex models in the future and (2) computationally cheaper for regularly sampled data. The Kalman filter is presented in appendix~\ref{sec:Kalman}. C-ARMA models can have an arbitrary number of autoregressive ($p$) and moving average ($q$) terms. In appendix~\ref{sec:ModelSelection}, we discuss information criteria that can be used to select the model order.

@@

The cause of perturbations, i.e. variability, in the flux emanating from an AGN is unknown, but we may construct a very general mathematical model for the form of the flux perturbations based on physical and statistical arguments. The region of space surrounding the central engine is vast. The inner edge of the accretion disk is thought to begin a few $r_{G} \sim 20$~AU ($M_{BH} = 10^{9} M_{\sun}$) away from the event horizon, while the outer edge is thought to lie at a distance of a few parsec from the event horizon. Contributions to the deviations in the flux most likely simultaneously arise from many individual locations in this system. Under the assumption that the distribution of energy released by each individual event is drawn from a distribution with finite second-moment, the central-limit-theorem assures us that the total energy introduced into the system at any instant will be drawn from a normal distribution i.e. some form of noise. Due to the non-linearity of the physics involved, it is quite likely that correlations exist between the energy released into the disk at different moments in time. Therefore, we shall model the correlations by treating the driving process $u$ in equation~\eqref{eq:LHS} as a correlated Gaussian noise process. In appendix~\ref{sec:RHS}, we argue why the natural mathematical form of the RHS should be given by
\begin{equation}\label{eq:RHS}
u = \beta_{0} (\mathrm{d}W) + \beta_{1} \mathrm{d}(\mathrm{d}W) + \ldots + \beta_{p-2} \mathrm{d}^{p-2}(\mathrm{d}W) + \beta_{p-1} \mathrm{d}^{p-1}(\mathrm{d}W).
\end{equation}
This form for the input stochastic impulses can be characterized by the power spectral density (PSD) of the impulses given by
\begin{equation}\label{eq:ImpulsePSD}
S_{uu}(\nu) = \frac{1}{2 \upi } |\beta(2 \upi \imath \nu)|^{2},
\end{equation}
where $\beta(z)$ is the moving average polynomial of equation~\eqref{eq:MACharPoly}. The PSD of the stochastic impulses tells about the timescales on which the stochastic impulses occur. Combining equation~\eqref{eq:LHS} with equation~\eqref{eq:RHS} yields the original C-ARMA process equation~\eqref{eq:CARMAIntro}. A C-ARMA process is a linear differential equation perturbed/driven by a random process. This driving process may be viewed as disturbances generated by MRI instabilities in the accretion disk. \textit{Thus we see that we can describe AGN luminosity variations in terms of a linear system driven by colored noise}.

@@

\begin{figure*}
    \includegraphics[width=\textwidth]{images/PowerOfSDSSK2.jpg}
    \caption{Why do AGN poorly-sampled AGN light curves look like a damped random walk (DRW)? We plot two PSD models - the DRW of \citet{Kelly09} in cyan and a damped harmonic oscillator (DHO) in orange. The DRW is a C-ARMA($1$,$0$) model i.e  the simplest possible C-ARMA model. The DHO is C-ARMA($2$,$q$) model i.e. it is a $2^{\text{nd}}$-order C-ARMA model. The vertical dashes are located at the time-intervals probed by NASA's \textit{K2} mission (purple), the SDSS survey (red), and an idealized combination of the two (green). On timescales longer than $\sim 10$~d, both PSDs are identical, explaining why the simpler DRW model is popular for light curves that do not probe short timescales. The availability of better data in the future will necessitate more sophisticated C-ARMA models.}
    \label{fig:PowerOfSDSSK2}
\end{figure*}

We shall discuss the C-ARMA($2$,$1$) process in detail because it has strong physical motivations and is relatively simple to interpret. Notice that in the process of linearising the original non-linear differential equation of equation~\eqref{eq:FluxDE}, we have not changed the order of the highest derivative of the flux that appears in that equation i.e. if the original equation governing the flux emitted by the accretion disk is of $p^{\mathrm{th}}$-order, the corresponding linearised C-ARMA process is also of $p^{\mathrm{th}}$-order. Most physical phenomenon are governed by $2^{\mathrm{nd}}$-order equations. Hence, we may expect to find that in the linear regime, AGN variability is a C-ARMA($2$,$q$) process with $q < 2$.

The general form of the C-ARMA($2$,$1$) process is
\begin{equation}\label{eq:CARMA21}
\mathrm{d}^{2}\chi + \alpha_{1} \mathrm{d}\chi + \alpha_{2} \chi = \beta_{0}(\mathrm{d}W) + \beta_{1} \mathrm{d}(\mathrm{d}W),
\end{equation}
where $[\alpha_{1}] = T^{-1}$, $[\alpha_{2}] = T^{-2}$, $[\beta_{0}] = [\chi]~T^{-3/2}$, and $[\beta_{1}] = [\chi]~T^{-1/2}$. The LHS determines how individual perturbations evolve over time. The RHS determines the shape of the driving process PSD. If $\beta_{1} = 0$, the driving noise process PSD is flat over all frequencies i.e. it is a white noise process. If $\beta \neq 0$, the disturbance process will have different amounts of power at different frequencies i.e. it is a colored' noise process. Equation~\eqref{eq:CARMA21} is of the same mathematical form as a \textit{Damped Harmonic Oscillator} (DHO) that is driven/perturbed by a colored noise process. Light curves that obey the C-ARMA($2$,$1$) process equation have perturbations from the mean flux level that individually behave like damped harmonic oscillators.
Given the identification of the  C-ARMA($2$,$1$) process with a DHO, $\alpha_{2} = \omega^{2}$ gives the angular frequency of the HO, and $\alpha_{1} = 2\omega\zeta_{\mathrm{dHO}}$ gives the damping ratio $\zeta_{\mathrm{dHO}}$ \citep[see chapter~7 for notation]{PanditWu}. The time-period of the harmonic oscillator is $T_{\mathrm{HO}} = 2 \upi/\omega$. The roots of the LHS of equation~\eqref{eq:CARMA21} are given by
\begin{equation}\label{eq:CARMA21Roots}
\rho_{1},\rho_{2} = -\omega\zeta_{\mathrm{dHO}} \pm \omega \sqrt{\zeta_{\mathrm{dHO}}^{2} - 1} = -\frac{\alpha_{1}}{2} \pm \sqrt{\frac{\alpha_{1}^{2}}{4} - \alpha_{2}}.
\end{equation}
The behaviour of the flux perturbations is non-physical in this context if $\zeta_{\mathrm{dHO}} = 0$, i.e. the flux perturbations are undamped and never die out. For $0 < \zeta_{\mathrm{dHO}} < 1$, the flux perturbations are under damped and will oscillate in value around the steady-state flux value with angular frequency $\omega' = \omega\sqrt{1 - \zeta_{\mathrm{dHO}}^{2}}$ but gradually decrease in amplitude. If $\zeta_{\mathrm{dHO}} = 1$ i.e. the perturbations are critically damped, flux perturbations will rapidly die down to the steady-state flux, while if $\zeta_{\mathrm{dHO}} > 1$, flux perturbations are over damped and exponentially decay to the steady-state flux. The rate of decay decreases with $\zeta_{\mathrm{dHO}}$, i.e. over damped flux perturbations with larger $\zeta$ take longer to decay, increasing the correlation timescale of the light curve. From equation~\eqref{eq:CARMAPSD}, the PSD of the C-ARMA($2$,$1$) process is given by
\begin{equation}\label{eq:CARMA21PSD}
S_{\chi\chi}(\nu) = \frac{1}{2\upi}\frac{\beta_{0}^{2} + 4\upi^{2}\beta_{1}^{2}\nu^{2}}{16\upi^{4}\nu^{4} + 4\upi^{2}(\alpha_{1}^{2}-2\alpha_{2})\nu^{2} + \alpha_{2}^{2}},
\end{equation} 
which has frequency dependent behaviour.

Figure~\ref{fig:PowerOfSDSSK2} shows the behavior of a DHO PSD with frequency. At very high- and low-frequencies, the DHO PSD looks like the damped random walk PSD of \citet{Kelly09} with $S_{\chi\chi} \sim 1/\nu^{2}$. There is a mid-range of frequencies where the $\nu^{4}$ term in the denominator dominates. At the lowest frequencies, the PSD is dominated by the frequency independent terms and asymptotically approachs $\beta_{2}^{2}/2\upi \alpha_{2}^{2}$. This frequency-dependent PSD behavior may explain why ground-based observations of AGN light curves suggest that AGN variability is well-modelled by a DRW whereas \textit{Kepler} observations of AGN suggest that the PSD is steeper than a DRW - the ground based observations are not sampling at high enough frequency to be able to see the $\nu^{4}$ dominated regime.

>>>>

% Abstract of the paper
\begin{abstract}
AGN exhibit rapid, high amplitude stochastic flux variability across the entire electromagnetic spectrum on timescales ranging from hours to years. The cause of this variability is poorly understood. We present a new method for using variability to (1) measure the time-scales on which flux perturbations evolve and (2) characterize the driving flux perturbations. We model the observed light curve of an AGN as a linear differential equation driven by stochastic impulses. Physically, the impulses could be local `hot-spots' in the accretion disk---the linear differential equation then governs how the hot spots evolve and dissipate. The impulse-response function of the accretion disk material is given by the Green's function of the linear differential equation. The timescales on which the hot-spots radiate energy is characterized by the powerspectrum of the driving stochastic impulses. We analyze the light curve of the \Kepler AGN Zw 229-15 and find that the observed variability behavior can be modeled as a damped harmonic oscillator perturbed by a colored noise process. The model powerspectrum turns over on time-scale $385$~d. On shorter time-scales, the log-powerspectrum slope varies between $2$ and $4$, explaining the behavior noted by previous studies. We recover and identify both the $5.6$~d and $67$~d timescales reported by previous work. These timescales represent the time-scale on which flux perturbations grow, and the time-scale on which flux perturbations decay back to the steady-state flux level respectively. We make the software package used to study light curves using our method, \textsc{k\={a}l\={i}}, available to the community.

%We have developed and tested a fast software package, \textsc{k\={a}l\={i}}, to analyze light curves using our method that we have made available . 
%We decompose the observed light curve fluctuations into stochastic impulses that drive an ordinary linear differential equation. The Green's function of the differential equation quantifies the time-domain evolution of perturbations to the emitted flux and hence probes dissipative physical processes in accretion flow. The PSD of the stochastic driving impulses characterizes the physical processes responsible for the flux perturbations.
%The PSD of the driving impulses is flat on long time-scales. On time-scales shorter than $1$~d, the impulse PSD begins to rise steeply possibly due to turbulent eddies in the accretion disk, or by the thermal motion of the accretion disk plasma.
\end{abstract}

@@

Numerous mechanisms have been posited for producing variability in accretion disks---see \citet{DoneAccretion}, \citet{MaccaroneAccretion}, and \citet{UttleyAccretion} for comprehensive overviews. While we are concerned primarily with accretion onto supermassive black holes (SMBH) in AGN, a fair fraction of the theory is developed with BHXRB accretion in mind. Although similarities exist between accretion in BHXRBs and AGN, the difference in energy densities is considerable, with AGNs having considerably higher luminosities and exhibiting variability on longer time-scales. For instance, quasi-periodic oscillations (QPO) are routinely observed in BHXRBs, whereas in AGN there few reports of QPOs \citep{GierlinskiQPO,Andrae13,GrindleyQPO}. While BHXRBs are commonly observed to shift between various spectral states, such transitions have not been observed in AGN to date \citep{Kelly11}.

AGN optical variability may arise due to fluctuations in the \citet{ShakuraSunyaev73} viscosity parameter $\alpha$. \citet{Lyubarskii97} showed that local stochastic fluctuations in the accretion disk are capable of producing the power-law power spectral densities (PSD) observed in AGN and BHXRBs. Such fluctuations are thought to propagate inwards over time suggesting that a powerful observational test is the observation of blue-lags \citep{UttleyAccretion}, i.e., light curve variations at shorter wavelengths lagging those at longer wavelengths. Evidence for the propagating fluctuation model was provided by \citet{Miyamoto88} who observed that hard X-ray variations in Cyg X-1 lagged behind soft X-ray variations with an inverse scaling between the time lag and the variability time-scale. \citet{Starling04} used the viscosity fluctuation model to impose lower limit of $\alpha \sim 0.01$ on the disk viscosity. A later study by \citet{King07} using AGN variability triggered by viscosity fluctuations suggest that $\alpha \sim 0.1$-$0.4$. \citet{Wood01} examined a variant of the \citet{Lyubarskii97} propagating fluctuations model in which episodic mass deposition at the outer edge of the accretion disk triggers viscosity fluctuations in the accretion disk. This model was revisited by \citet{Titarchuk07} who generalized the treatment of \citet{Wood01} to include time-dependent fluctuation sources located randomly on the disk face. \citet{Titarchuk07} follow the evolution of perturbations to the mass surface density and luminosity arising from viscosity that is a power-law function of disk radius. Subsequently, \citet{Kelly11} studied solutions to the surface mass density equation perturbed by stochastic fluctuations and showed that the resulting light curve can be expressed as a linear combination of multiple Ornstein-Uhlenbeck processes \citep{Gillespie96,Kelly09}.

@@

Studies of AGN variability model the observed light curve as stochastic processes i.e mathematically the light curve as a function of time is a random variable that correlates with previous values of the same light curve. Prescriptions for stochastic models require two properties to be specified: 1. the distribution from which to draw the random values of the light curve; and 2. the strength of the correlation of a given value of the light curve with previous values of the same light curve. Most studies of AGN variability (implicitly) invoke some form of the central limit theorem to argue that property 1 should be a Gaussian distribution which then lets them specify property 2 using a second-order statistic such as the power spectral density (PSD), or the auto-covariance function (ACVF), or the structure function (SF). The most straightforward method of fitting such a stochastic model to a light curve is by direct estimation of the PSD of the observed light curve. Using periodogram estimators of the PSD, the observed PSD can be compared to the predicted theoretical PSD as advocated in \citet*{Uttley02}. Alternatively, it is also possible to estimate the ACVF of the observed light curve and fit it to the theoretical ACVF of a given stochastic process. Equivalently, following \citet*{Kasliwal15}, estimates of the structure function (SF) of the observed light curve can be fit to the theoretical SF computed either directly from the theoretical ACVF, or from Monte-Carlo simulations of light curves generated using the prescribed model PSD.

@@

Recently, \citet{Kelly14} suggested that stochastic variability in a large class of astrophysical sources may be modelled using Continuous-time AutoRegressive Moving Average (C-ARMA) processes. The attractiveness of using C-ARMA processes to model stochastic phenomenon stems from two factors: 1. C-ARMA processes directly specify property 2, i.e., the correlation structure of the light curve by providing a differential equation that governs how the light curve evolves over time. This alternative to providing a second-order statistic is powerful because it makes the light curve evolution more amenable to physical interpretation as is shown in this work. It also provides some idea of how to replace Gaussian distributions with more generalized $\alpha$-stable distributions that may be truer to reality \citep{BrockwellMarquardt05, BrockwellLinder09}; and 2. flexibility of the shape of their PSD. Since the PSD of a C-ARMA process is a rational function, it can be made arbitrarily complex to model a very wide range of phenomenon. This flexibility is achieved by allowing C-ARMA processes to have multiple Autoregressive (quantified by $p$) and Moving Average (quantified by $q$) terms under the constraint $q < p$. If $\chi$ is a small flux perturbation from the mean flux level, then a C-ARMA($p$,$q$) process that models the time evolution of $\chi$ has the general form
\begin{multline}\label{eq:CARMAIntro}
\mathrm{d}^{p}\chi + \alpha_{1} \mathrm{d}^{p-1}\chi + \ldots + \alpha_{p-1} \mathrm{d}\chi + \alpha_{p} \chi = \\ \beta_{0} (\mathrm{d}W) + \beta_{1} \mathrm{d}(\mathrm{d}W) + \ldots + \beta_{p-1} \mathrm{d}^{p-1}(\mathrm{d}W)
\end{multline}
where $W$ is a \textit{Wiener} process and is responsible for introducing stochasticity into the behavior of $\chi$. The $\alpha$ and $\beta$ are constants that govern the behavior of the resulting process and have dimensions of various powers of time. In this paper, we demonstrate how C-ARMA processes can arise from the physical processes in accretion disks and present a Green's function-based method for understanding what a C-ARMA process tells us about accretion physics.

@@

Appendices~\ref{sec:Fitting} through~\ref{sec:ModelSelection} present details of inferring the properties of the C-ARMA process from noisy observations using the Kalman filter. Alternatives exist to using the Kalman filter for inferring the parameters of a C-ARMA process. We discuss why the Kalman filter is the optimal choice of fitting techniques in appendix~\ref{sec:Fitting}
%In brief, the Kalman filter is both statistically optimal and computationally efficient. In order to use the Kalman filter, it is necessary to cast the C-ARMA process into \textit{state-space} form. 
and suggest an alternative to the traditional \textit{controllable canonical form} state-space representation used by \citet{JonesAckerson90} and \citet{Brockwell14}. Our choice of state-space representations is (1) better suited for extension to more complex models in the future and (2) computationally cheaper for regularly sampled data. The Kalman filter is presented in appendix~\ref{sec:Kalman}. C-ARMA models can have an arbitrary number of autoregressive ($p$) and moving average ($q$) terms. In appendix~\ref{sec:ModelSelection}, we discuss information criteria that can be used to select the model order.

@@

The cause of perturbations, i.e., variability, in the flux emanating from an AGN is unknown, but we may construct a very general mathematical model for the form of the flux perturbations based on physical and statistical arguments. The region of space surrounding the central engine is vast. The inner edge of the accretion disk is thought to begin a few $r_{G} \sim 20$~AU ($M_{BH} = 10^{9} M_{\sun}$) away from the event horizon, while the outer edge is thought to lie at a distance of a few parsec from the event horizon. Contributions to the deviations in the flux most likely simultaneously arise from many individual locations in this system. Under the assumption that the distribution of energy released by each individual event is drawn from a distribution with finite second-moment, the central-limit-theorem assures us that the total energy introduced into the system at any instant will be drawn from a normal distribution, i.e., some form of noise. Due to the non-linearity of the physics involved, it is quite likely that correlations exist between the energy released into the disk at different moments in time. Therefore, we shall model the correlations by treating the driving process $u$ in equation~\eqref{eq:LHS} as a correlated Gaussian noise process. In appendix~\ref{sec:RHS}, we argue why the natural mathematical form of the RHS should be given by
\begin{equation}\label{eq:RHS}
u = \beta_{0} (\mathrm{d}W) + \beta_{1} \mathrm{d}(\mathrm{d}W) + \ldots + \beta_{p-2} \mathrm{d}^{p-2}(\mathrm{d}W) + \beta_{p-1} \mathrm{d}^{p-1}(\mathrm{d}W).
\end{equation}
This form for the input stochastic impulses can be characterized by the power spectral density (PSD) of the impulses given by
\begin{equation}\label{eq:ImpulsePSD}
S_{uu}(\nu) = \frac{1}{2 \upi } |\beta(2 \upi \imath \nu)|^{2},
\end{equation}
where $\beta(z)$ is the moving average polynomial of equation~\eqref{eq:MACharPoly}. The PSD of the stochastic impulses tells about the timescales on which the stochastic impulses occur. Combining equation~\eqref{eq:LHS} with equation~\eqref{eq:RHS} yields the original C-ARMA process equation~\eqref{eq:CARMAIntro}. A C-ARMA process is a linear differential equation perturbed/driven by a random process. This driving process may be viewed as disturbances generated by MRI instabilities in the accretion disk. \textit{Thus we see that we can describe AGN luminosity variations in terms of a linear system driven by colored noise}.

@@

\begin{figure*}
    \includegraphics[width=\textwidth]{images/PowerOfSDSSK2.jpg}
    \caption{Why do AGN poorly-sampled AGN light curves look like a damped random walk (DRW)? We plot two PSD models - the DRW of \citet{Kelly09} in cyan and a damped harmonic oscillator (DHO) in orange. The DRW is a C-ARMA($1$,$0$) model i.e  the simplest possible C-ARMA model. The DHO is C-ARMA($2$,$q$) model, i.e., it is a $2^{\text{nd}}$-order C-ARMA model. The vertical dashes are located at the time-intervals probed by NASA's \textit{K2} mission (purple), the SDSS survey (red), and an idealized combination of the two (green). On timescales longer than $\sim 10$~d, both PSDs are identical, explaining why the simpler DRW model is popular for light curves that do not probe short timescales. The availability of better data in the future will necessitate more sophisticated C-ARMA models.}
    \label{fig:PowerOfSDSSK2}
\end{figure*}

We shall discuss the C-ARMA($2$,$1$) process in detail because it has strong physical motivations and is relatively simple to interpret. Notice that in the process of linearising the original non-linear differential equation of equation~\eqref{eq:FluxDE}, we have not changed the order of the highest derivative of the flux that appears in that equation, i.e., if the original equation governing the flux emitted by the accretion disk is of $p^{\mathrm{th}}$-order, the corresponding linearised C-ARMA process is also of $p^{\mathrm{th}}$-order. Most physical phenomenon are governed by $2^{\mathrm{nd}}$-order equations. Hence, we may expect to find that in the linear regime, AGN variability is a C-ARMA($2$,$q$) process with $q < 2$.

The general form of the C-ARMA($2$,$1$) process is
\begin{equation}\label{eq:CARMA21}
\mathrm{d}^{2}\chi + \alpha_{1} \mathrm{d}\chi + \alpha_{2} \chi = \beta_{0}(\mathrm{d}W) + \beta_{1} \mathrm{d}(\mathrm{d}W),
\end{equation}
where $[\alpha_{1}] = T^{-1}$, $[\alpha_{2}] = T^{-2}$, $[\beta_{0}] = [\chi]~T^{-3/2}$, and $[\beta_{1}] = [\chi]~T^{-1/2}$. The LHS determines how individual perturbations evolve over time. The RHS determines the shape of the driving process PSD. If $\beta_{1} = 0$, the driving noise process PSD is flat over all frequencies, i.e., it is a white noise process. If $\beta \neq 0$, the disturbance process will have different amounts of power at different frequencies, i.e., it is a `colored' noise process. Equation~\eqref{eq:CARMA21} is of the same mathematical form as a \textit{Damped Harmonic Oscillator} (DHO) that is driven/perturbed by a colored noise process. Light curves that obey the C-ARMA($2$,$1$) process equation have perturbations from the mean flux level that individually behave like damped harmonic oscillators.
Given the identification of the  C-ARMA($2$,$1$) process with a DHO, $\alpha_{2} = \omega^{2}$ gives the angular frequency of the HO, and $\alpha_{1} = 2\omega\zeta_{\mathrm{dHO}}$ gives the damping ratio $\zeta_{\mathrm{dHO}}$ \citep[see chapter~7 for notation]{PanditWu}. The time-period of the harmonic oscillator is $T_{\mathrm{HO}} = 2 \upi/\omega$. The roots of the LHS of equation~\eqref{eq:CARMA21} are given by
\begin{equation}\label{eq:CARMA21Roots}
\rho_{1},\rho_{2} = -\omega\zeta_{\mathrm{dHO}} \pm \omega \sqrt{\zeta_{\mathrm{dHO}}^{2} - 1} = -\frac{\alpha_{1}}{2} \pm \sqrt{\frac{\alpha_{1}^{2}}{4} - \alpha_{2}}.
\end{equation}
The behaviour of the flux perturbations is non-physical in this context if $\zeta_{\mathrm{dHO}} = 0$, i.e., the flux perturbations are undamped and never die out. For $0 < \zeta_{\mathrm{dHO}} < 1$, the flux perturbations are under damped and will oscillate in value around the steady-state flux value with angular frequency $\omega' = \omega\sqrt{1 - \zeta_{\mathrm{dHO}}^{2}}$ but gradually decrease in amplitude. If $\zeta_{\mathrm{dHO}} = 1$, i.e., the perturbations are critically damped, flux perturbations will rapidly die down to the steady-state flux, while if $\zeta_{\mathrm{dHO}} > 1$, flux perturbations are over damped and exponentially decay to the steady-state flux. The rate of decay decreases with $\zeta_{\mathrm{dHO}}$, i.e., over damped flux perturbations with larger $\zeta$ take longer to decay, increasing the correlation timescale of the light curve. From equation~\eqref{eq:CARMAPSD}, the PSD of the C-ARMA($2$,$1$) process is given by
\begin{equation}\label{eq:CARMA21PSD}
S_{\chi\chi}(\nu) = \frac{1}{2\upi}\frac{\beta_{0}^{2} + 4\upi^{2}\beta_{1}^{2}\nu^{2}}{16\upi^{4}\nu^{4} + 4\upi^{2}(\alpha_{1}^{2}-2\alpha_{2})\nu^{2} + \alpha_{2}^{2}},
\end{equation} 
which has frequency-dependent behaviour.

Figure~\ref{fig:PowerOfSDSSK2} shows the dependence of a DHO PSD on frequency. At very high- and low-frequencies, the DHO PSD looks like the damped random walk PSD of \citet{Kelly09} with $S_{\chi\chi} \sim 1/\nu^{2}$. There is a mid-range of frequencies where the $\nu^{4}$ term in the denominator dominates. At the lowest frequencies, the PSD is dominated by the frequency independent terms and asymptotically approachs $\beta_{2}^{2}/2\upi \alpha_{2}^{2}$. This frequency-dependent PSD behavior may explain why ground-based observations of AGN light curves suggest that AGN variability is well-modelled by a DRW whereas \textit{Kepler} observations of AGN suggest that the PSD is steeper than a DRW - the ground based observations are not sampling at high enough frequency to be able to see the $\nu^{4}$ dominated regime.

Individual accretion disk perturbations are smeared by processes such as turbulent dissipation, rotation and thermal dissipation. The nature of these processes can be inferred via the Green's function approach as discussed in section~\ref{sec:Meaning}. The Green's function of the DHO is given by
\begin{equation}\label{eq:DHOGFunc}
G(t) = \frac{\mathrm{e}^{\rho_{1}t} - \mathrm{e}^{\rho_{2}t}}{\rho_{1} - \rho_{2}},
\end{equation}
where $\rho_{1}$ and $\rho_{2}$ are the roots of the autoregressive characteristic polynomial of equation~\eqref{eq:CARMA21Roots}. If $\zeta_{\mathrm{dHO}} > 1$, i.e., flux perturbations are constrained to be damped out by the dissipative processes in the accretion disk, then the roots are purely real-valued. A perturbation to the accretion disk flux in the form of an impulse at $t = t_{0}$ will result in the flux increasing from $t_{0}$ to $t_{max}$ given by
\begin{equation}\label{eq:maxT}
t_{\mathrm{max}} = \frac{1}{\rho_{2} - \rho_{1}} \ln \left| \frac{\rho_{1}}{\rho_{2}} \right|.
\end{equation}
After $t_{\mathrm{max}}$, the effect of the original perturbation decays to the steady-state flux level. Although the decay is not a pure exponential, it is useful to characterize the decay rate by an e-folding time-scale $t_{\mathrm{e-fold}}$. It may be computed by solving
\begin{equation}\label{eq:TEFold}
G(t_{\mathrm{e-fold}}) = \frac{G(t_{\mathrm{max}})}{\mathrm{e}},
\end{equation} 
numerically for $t_{\mathrm{e-fold}}$.
